---
title: "Hypothesis Testing"
author: "Alicia Halhed"
date: "04/03/2020"
output: word_document
---
## Set up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# local
setwd("~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/R-env")
# or on graham.sharcnet.ca
# setwd("/home/ahalhed/red-squirrel-w2020")
```

```{r}
library(qiime2R)
library(tidyverse)
library(phyloseq)
library(ggpubr)
```

## Phyloseq

```{r}
# this was run on SHARCNET and the environment was downloaded locally due to high memory requirement
ps <- qza_to_phyloseq(features = "~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/filtered-table.qza",
                      tree = "~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/trees/rooted_tree.qza",
                      taxonomy = "~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/taxonomy/GG-taxonomy.qza",
                      metadata = "RS_meta.tsv")
```

```{r}
# this computes the following: c("Observed","Chao1","ACE","Shannon","Simpson","InvSimpson","Fisher")
# have "Observed" and "Shannon" from QIIME2 as well (hopefully these will be the same)
ps_richness <- estimate_richness(ps) # took a few minutes to run
```
```{r}
# load in additional values from QIIME2
q2.shannonp <- read_qza("~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/core-metrics-phylogenetic/shannon_vector.qza")$data
q2.shannonnp <- read_qza("~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/core-metrics-non-phylogenetic/shannon_vector.qza")$data
q2.faith_pd <- read_qza("~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/core-metrics-phylogenetic/faith_pd_vector.qza")$data
q2.evenness <- read_qza("~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/core-metrics-phylogenetic/evenness_vector.qza")$data
q2.otus <- read_qza("~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/core-metrics-non-phylogenetic/observed_otus_vector.qza")$data
merge(q2.evenness, q2.shannonnp, q2.shannonp, q2.faith_pd, q2.otus, by=0, all=T)
```

```{r}
# append additional values from QIIME2 to the above richness data frame
merged_richness <- merge(q2.evenness, q2.shannonnp, by=0, all = T) %>%
  column_to_rownames(var = "Row.names") %>%
  rename("q2.npshannon" = "shannon") %>%
  merge(., q2.shannonp, by=0, all = T) %>%
  column_to_rownames(var = "Row.names") %>%
  merge(., q2.faith_pd, by=0, all = T) %>%
  column_to_rownames(var = "Row.names") %>%
  merge(., q2.otus, by=0, all = T) %>%
  column_to_rownames(var = "Row.names") %>%
  rename("q2.pielou_e" = "pielou_e",
         "q2.pshannon" = "shannon",
         "q2.faith_pd" = "faith_pd",
         "q2.observed_otus" = "observed_otus") %>%
  merge(., ps_richness, by=0, all = T) %>%
  column_to_rownames(var = "Row.names") %>%
  select("q2.pshannon", "q2.npshannon", "Shannon",
         "q2.observed_otus", "Observed", everything())
View(merged_richness)
```

```{r}
# remove R objects no longer needed
rm(q2.evenness, q2.faith_pd, q2.otus, q2.shannonnp, q2.shannonp)
```

## Hypothesis 1 - Environment
Group individuals by grid/location
```{r}

```

## Hypothesis 2 - Diet
Group individuals by food supplementation
```{r}

```

## Hypothesis 3 - Parental care
Group individuals by DamID and SireID
```{r}

```


## Hypothesis 4 - Dispersal
Group individuals by number of different locations sampled.

Maybe something with comparing first and last location sampled. Account for time difference?

```{r}
# GET the metadata
rs_q2_metadata <- read.table("~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/RS_meta.tsv", sep="\t")
colnames(rs_q2_metadata) <- c("SampleID", "Grid", "Location X", "Location Y", "Sex", "Age", "Month", "Season", "Year", "Squirrel.ID", "SireID", "DamID", "CollectionDate", "FoodSupplement", "BirthYear", "Location", "Date")
# add dispersal numbers onto the metadata
dispersal_metadata <- rs_q2_metadata %>%
  group_by(Squirrel.ID) %>% 
  mutate(CollectionDate = str_remove_all(CollectionDate, "-"),
         Unique.Locations = n_distinct(Location))
# take only the most recently collected ones
rs_sub_meta <- rs_q2_metadata %>%
  group_by(Squirrel.ID) %>% 
  mutate(CollectionDate = str_remove_all(CollectionDate, "-"),
         Unique.Locations = n_distinct(Location)) %>%
  top_n(1, CollectionDate)
```
```{r}
# bring in distance matrices
dm_phylo_bray <- read_qza("../core-metrics-phylogenetic/bray_curtis_distance_matrix.qza")
dm_phylo_jaccard <- read_qza("../core-metrics-phylogenetic/jaccard_distance_matrix.qza")
```
### Jaccard plot
The original paper compared the Jaccard distance between related individuals. I am doing that with the number of sampling locations.
```{r}
df_jaccar_loc <- dm_phylo_jaccard$data %>% 
  as.matrix %>% as.data.frame %>%
  rownames_to_column(var = "SampleID1") %>%
  pivot_longer(-SampleID1, names_to = "SampleID2", values_to = "Distance") %>%
  unique() %>% # this removes the duplications of the data
  left_join(rs_q2_metadata %>% 
              group_by(Squirrel.ID) %>% 
              mutate(Unique.Locations = n_distinct(Location)),
            by = c("SampleID1" = "SampleID")) %>%
  select(SampleID1, Unique.Locations, SampleID2, Distance) %>%
  rename("Unique.Locations1"="Unique.Locations") %>%
  left_join(rs_q2_metadata %>% 
              group_by(Squirrel.ID) %>% 
              mutate(Unique.Locations = n_distinct(Location)),
            by = c("SampleID2" = "SampleID")) %>%
  select(SampleID1, Unique.Locations1, SampleID2, Unique.Locations, Distance) %>%
  rename("Unique.Locations2"="Unique.Locations") %>%
  unite(Pairs, c(Unique.Locations1, Unique.Locations2), remove=FALSE, sep = "-") %>%
  subset(SampleID1 %in% rs_sub_meta$SampleID) %>%
  subset(SampleID2 %in% rs_sub_meta$SampleID)
View(df_jaccar_loc)
```
```{r}
# mean pairs to compare
comparisons <- list(c("1-1", "1-2"), c("1-1", "1-3"), 
                    c("1-1", "1-4"), c("1-1", "1-5"), 
                    c("1-1", "1-6"), c("1-1", "1-7"), 
                    c("1-1", "1-8"), c("1-1", "1-9"))
# plot
df_jaccar_loc %>%
  filter(Distance != 0) %>%
  filter(Unique.Locations1 == 1) %>%
  ggplot(aes(x=Pairs, y=Distance, group=Pairs)) + 
  geom_boxplot() + theme_bw()  +
  xlab("Pairs Compared") +
  ylab("Jaccard Distance") +
  stat_compare_means(comparisons = comparisons,
                     symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), 
                                        symbols = c("****", "***", "**", "*", "ns"))) +
  stat_compare_means(label.y = 1.2) +
  stat_summary(fun.y=mean, geom="point", shape=20, size=1, color="red", fill="red")
```
### Some stats
```{r}
# full data (all samples)
vegan::anosim(dm_phylo_bray$data, dispersal_metadata$Unique.Locations, permutations = 1000)
# only the most recently collected
# subset a distance matrix?
vegan::anosim(dm_phylo_bray$data, rs_sub_meta$Unique.Locations, permutations = 1000)
```
```{r}
# full data (all samples)
vegan::adonis(dm_phylo_bray$data ~ Unique.Locations*CollectionDate, 
              data = dispersal_metadata, 
              permutations = 1000)
# only the most recently collected
vegan::adonis(dm_phylo_bray$data ~ Unique.Locations*CollectionDate, 
              data = rs_sub_meta, 
              permutations = 1000)
```


