---
title: "Red Squirrel Proposal Hypotheses"
author: "Alicia Halhed"
date: "20/03/2020"
output: word_document
---

## Set Up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# set working directory
setwd("~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/R-env")
# attach required packages
library(tidyverse)
library(qiime2R)
```

```{r}
# get the metadata
rs_q2_metadata <- read.table("~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/RS_meta.tsv", sep="\t")
```

## Proposal Figure 3
```{r}
# load in PCoA
rs_PCoA_jaccard <- read_qza("~/OneDrive - University of Guelph/Alicia's Thesis/red-squirrel-w2020/core-metrics-phylogenetic/jaccard_pcoa_results.qza")
```

### Panel 3-1 
Ordination of grids
```{r}
# combine PCoA with metadata
rs_PCoA_V <- left_join(rs_PCoA_jaccard$data$Vectors, rs_q2_metadata)
```

```{r}
# axes 1 and 2
rs_PCoA_plot <- ggplot(rs_PCoA_V, 
                       aes(x=PC1, y=PC2, col = Grid, group = Squirrel.ID)) + 
  geom_point() + # points are samples
  theme_bw()
rs_PCoA_plot
# add individual lines
rs_PCoA_plot + geom_line() # lines are individuals
# add ellipses by grid
rs_PCoA_plot + stat_ellipse(aes(group=Grid))
# add ellipses by grid and individual lines
# panel by year
rs_PCoA_plot + 
  geom_line() + 
  stat_ellipse(aes(group=Grid)) +
  facet_grid(Season~Year)

```

### Panel 3-2 
Ordination of grids with diet

### Panel 3-3 
Ordination of grids with parent-offspring relationships

### Panel 3-4
Ordination of grids with movement between grids

## Proposal Figure 4

### Left Panel
Boxplot of dissimilarity of shared diets

### Right Panel
Boxplot of dissimilarity of parent-offspring pairs vs whole grid


## Proposal Figure 5
Line graph of dissimilarity between the focal individual and others sampled from the same location over time
I am going to focus on individuals sampled at two locations (ideally multiple times at both locations)